Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase Retrieval with Automatic Differentiation

### 2022-12-12 src/tex/ms.tex
[showyourwork] overleaf sync

In each case, we face serious limitations from an imperfect knowledge of the \ac{prf} -- the map of the intra-pixel and inter-pixel variations in detector sensitivity; and also of the \ac{psf} -- the diffraction-limited pattern by which light from a point source like a starspreads across a detector.

@@

\begin{figure*}
    \centering
    % \includegraphics[width=\textwidth]{figures/optics.pdf}
    \includegraphics[width=\textwidth]{figs/optics.pdf}
    \caption{Summary of the optical configuration. Left panel: The top plot show the \ac{opd} of the pupil at the aperture of the telescope. The binary values create a half-wave step at the mean observation wavelength. The bottom plot shows the resulting large single-star \ac{psf} without aberrations applied. Middle panel: The top plot shows the total \ac{opd} of the pupil with the optical aberrations applied using low-order Zernike polynomials that is used to generate the data. The bottom plot shows the resulting single-star \ac{psf} with the aberrations applied. Clearly these aberrations have a large effect on the \ac{psf} and would make recovering information very difficult without appropriate calibration. Right panel: The top plot shows the histogram of the PRF that is applied in the focal plane. These values have a large spread and would greatly affect any results without calibration. The bottom plot shows the full PRF across the whole detector, with a small zoomed region used to show the fine detail that can not be seen when examining the full detector.}
    \label{fig:optics}
    % \script{gen_optics.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figs/data_resid.pdf}
    \caption{Left: single image of the full four that is used to create the full data set. This image has had both photon and detector noise applied. By eye it is clearly difficult to disentangle the astrophysical information. The large number of overlapping \ac{psf}s is chosen in order to spread light across the majority of the detector so that we encode the PRF information for as much of the detector as possible. Middle: the residual of this single image from the data and the initial uncalibrated model. Clearly these residuals are large, showing that there is a large amount of calibration required. Right: this same residual after the model has been optimised. The residual values are much smaller and at the noise floor. A small zoomed region is shown so that the individual pixel-level residuals can be seen.}
    \label{fig:data_residual}
    % \script{gen_data_resid.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    % \includegraphics[width=0.8\textwidth]{figures/astro_params.pdf}
    \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
    \caption{Left: recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds. The bottom section shows the individual residuals for each. Clearly these parameters have been well recovered through the optimisation. Right: the recovery of the flux parameters in units of photons. The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals. The error bars are the 1-$\sigma$ deviations calculated from the covariance Hessian matrix marginalised over positions, fluxes and aberrations.}
    \label{fig:astro_params}
    % \script{gen_astro_params.py}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figs/aberrations.pdf}
    \caption{Recovery of the optical aberrations after optimisation. Left: the amplitudes of the Zernike polynomial coefficients, with the top section showing the correlation plot of the true and recovered values, and the bottom showing the individual residuals. Middle and right: the true total \ac{opd} and the residuals generated by these values respectively. The error bars are the 1-$\sigma$ deviations calculated from the covariance Hessian matrix marginalised over positions, fluxes and aberrations.}
    \label{fig:aberrations}
    % \script{gen_aberrations.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{figs/divergence.pdf}
    \caption{Parameter recovery of a single star over varying \ac{snr}. Left: Absolute positional residuals vs total flux. Right: Fractional flux error vs total flux.}
    \label{fig:diverge}
    % \script{gen_FF.py}
\end{figure*}

>>>>

In each case, we face serious limitations from an imperfect knowledge of the \ac{prf} -- the map of the intra-pixel and inter-pixel variations in detector sensitivity; and also of the \ac{psf} -- the diffraction-limited pattern by which light from a point source like a star spreads across a detector.

@@

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/optics.pdf}
    % \includegraphics[width=\textwidth]{figs/optics.pdf}
    \caption{Summary of the optical configuration. Left panel: The top plot show the \ac{opd} of the pupil at the aperture of the telescope. The binary values create a half-wave step at the mean observation wavelength. The bottom plot shows the resulting large single-star \ac{psf} without aberrations applied. Middle panel: The top plot shows the total \ac{opd} of the pupil with the optical aberrations applied using low-order Zernike polynomials that is used to generate the data. The bottom plot shows the resulting single-star \ac{psf} with the aberrations applied. Clearly these aberrations have a large effect on the \ac{psf} and would make recovering information very difficult without appropriate calibration. Right panel: The top plot shows the histogram of the PRF that is applied in the focal plane. These values have a large spread and would greatly affect any results without calibration. The bottom plot shows the full PRF across the whole detector, with a small zoomed region used to show the fine detail that can not be seen when examining the full detector.}
    \label{fig:optics}
    \script{plot_optics.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    % \includegraphics[width=\textwidth]{figs/data_resid.pdf}
    \includegraphics[width=\textwidth]{figures/data_resid.pdf}
    \caption{Left: single image of the full four that is used to create the full data set. This image has had both photon and detector noise applied. By eye it is clearly difficult to disentangle the astrophysical information. The large number of overlapping \ac{psf}s is chosen in order to spread light across the majority of the detector so that we encode the PRF information for as much of the detector as possible. Middle: the residual of this single image from the data and the initial uncalibrated model. Clearly these residuals are large, showing that there is a large amount of calibration required. Right: this same residual after the model has been optimised. The residual values are much smaller and at the noise floor. A small zoomed region is shown so that the individual pixel-level residuals can be seen.}
    \label{fig:data_residual}
    \script{plot_data_resid.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/astro_params.pdf}
    % \includegraphics[width=0.8\textwidth]{figs/astro_params.pdf}
    \caption{Left: recovery of the individual RA-Dec positions of each star, with the top section showing the correlation between the true and recovered values in units of arcseconds. The bottom section shows the individual residuals for each. Clearly these parameters have been well recovered through the optimisation. Right: the recovery of the flux parameters in units of photons. The top section shows the correlation between the true and recovered values, and the bottom showing the resulting residuals. The error bars are the 1-$\sigma$ deviations calculated from the covariance Hessian matrix marginalised over positions, fluxes and aberrations.}
    \label{fig:astro_params}
    \script{plot_astro_params.py}
\end{figure*}

\begin{figure*}
    \centering
    \includegraphics[width=\textwidth]{figures/aberrations.pdf}
    \caption{Recovery of the optical aberrations after optimisation. Left: the amplitudes of the Zernike polynomial coefficients, with the top section showing the correlation plot of the true and recovered values, and the bottom showing the individual residuals. Middle and right: the true total \ac{opd} and the residuals generated by these values respectively. The error bars are the 1-$\sigma$ deviations calculated from the covariance Hessian matrix marginalised over positions, fluxes and aberrations.}
    \label{fig:aberrations}
    \script{plot_aberrations.py}
\end{figure*}

@@

\begin{figure*}
    \centering
    % \includegraphics[width=0.8\textwidth]{figs/divergence.pdf}
    \includegraphics[width=0.8\textwidth]{figures/divergence.pdf}
    \caption{Parameter recovery of a single star over varying \ac{snr}. Left: Absolute positional residuals vs total flux. Right: Fractional flux error vs total flux.}
    \label{fig:diverge}
    \script{plot_divergence.py}
\end{figure*}
